### 2020-05-21 ArticleAcortado.tex
Introduccion y conclusiones

@@

\theoremstyle{definition}
\newtheorem{proposition}{Proposition}[section]
\newtheorem{fact}{Fact}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{obs}{Observation}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{propdef}{Proposition / Definition}[section]
\newtheorem{remark}{Remark}[section]

\newcommand{\cc}{\mathfrak{c}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Hc}{\mathcal{H}}
\newcommand{\Lan}{\mathcal{L}}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\newcommand{\clist}{\mathfrak{c}_{1}, \cdots, \mathfrak{c}_m}
\newcommand{\morph}[1]{\sim_#1}
\newcommand{\vlst}[2]{#1_1,\dots, #1_{#2}}
\newcommand{\gnp}{G(n,\beta_1/n^{a_1-1}, \dots,\beta_l/n^{a_l-1})}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\ehr}{\textsc{Ehr}}
\newcommand{\PR}[1]{\mathrm{Pr}\big(#1\big)}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\InR}[1]{\left\{ #1_R \right\}_{R\in \sigma}}
\newcommand{\sep}{\noindent\rule{2cm}{0.4pt}}
\newcommand{\aut}{\mathrm{aut}}

@@

\todo[inline]{INCOMPLETA, PENDIENTE DE SER REVISADA}
Since the work of Erd\"os and R\`enyi on the evolution of random graphs
\cite{erdHos1960evolution} the study of the asymptotic properties of random
structures has played a relevant role in combinatorics and computer science.
A central theme in this topic is, given a succession $(G_n)_n$ of random
structures of some sort and a property $P$, to determine the limit probability
that $G_n$ satisfies $P$ or to determine whether that limit exists. \par
One approach that has proven to be useful is to classify the properties $P$
according to the logical languages they can be defined in. We say that the 
succession $(G_n)_n$ obeys a convergence law with respect to some logical language
$\mathcal{L}$ if for any given property $P$ expressible in $\mathcal{L}$ the 
probability that $G_n$ satisfies $P$ tends to some limit as $n$ grows to infinity.
We say that $(G_n)_n$ obeys a zero-one law with respect to $\mathcal{L}$ if
that limit is always either zero or one. 
The seminal theorem on this topic, due to Fagin \cite{fagin1976probabilities}
and Glebskii et al. 
\cite{glebskii1969range} independently, states that if $G_n$ denotes a labeled
graph with $n$ vertices picked uniformly at random among all $2^{\binom{n}{2}}$ 
possible then $(G_n)_n$ satisfies a zero-one law with respect to the first order
(FO) language of graphs. \par
Originally this result was proven in the broader context of relational
structures but it was in the theory of random graphs were the study of 
other zero-one and convergence laws became more prominent. In particular,
the asymptotic behavior of FO logic in the binomial model of random graphs 
$G(n,p)$ has been extensively studied. In this model, introduced by Gilbert 
\cite{gilbert1959random}, a random graph is obtained from $n$ labeled vertices
by adding each possible edge with probability $p$ independently. When $p=1/2$
this distribution of random graphs coincides with the uniform one, mentioned 
above. In general, for the case where $p$ is a constant probability a slight 
generalization of the proofs in \cite{fagin1976probabilities} and 
\cite{glebskii1969range} works and $G(n,p)$ satisfies a zero-one law 
for FO logic. If we consider $p(n)$ a decreasing function of the form
$n^{-\alpha}$ we can ask the question of what are the values of $\alpha$
for which $G(n,p(n))$ obeys a zero-one or a convergence law for FO logic. 
In \cite{shelah1988zero} Shelah and Spencer gave a complete answer 
for the range $\alpha\in (0,1)$. Among other results, they proved that
if $\alpha$ is an irrational number in this interval then
$G(n,p(n))$ obeys a zero-one law for FO logic, while if $\alpha$
is a rational number in the same range then $G(n,p(n))$ does not
even satisfy a convergence law for FO logic. The case $\alpha=1$
was later solved by Lynch in \cite{lynch1992probabilities}. A weaker
form of the main theorem in that article states the following:
\begin{theorem} 
	For any FO sentence $\phi$, the function
	$F_\phi: (0,\infty)\rightarrow [0,1]$ given by 
	\[ F_\phi(\beta) = \Ln \mathrm{Pr}\big( \mathrm{G}(n,\beta/n)
	\text{ satisfies } \phi   \big) \]
	is well defined and analytic. In particular, for any
	$\beta \geq 0$ the model $\mathrm{G}(n,\beta/n)$
	obeys a convergence law for FO logic. 
\end{theorem}

The analyticity of these asymptotic probabilities with respect to 
the parameter $\beta$ implies that FO properties cannot "capture" sudden
changes that occur in the random graph $\mathrm{G}(n,\beta/n)$
as $\beta$ changes. Given $p(n)$ a probability, $P$ a property of graphs,
and $Q$ a sufficient condition for $P$ - i.e., a property that 
implies $P$ -, we say that $Q$ explains $P$ if $\mathrm{G}(n,p(n))$ satisfies 
the converse implication $P \implies Q$ asymptotically almost surely
(a.a.s.). A notable example of this phenomenon happens in the range 
$p(n)= \log(n)/n + \beta/n$ with $\beta$ constant. Erd\"os and R\`enyi
\cite{erdHos1960evolution} 
showed that for probabilities of this form $\mathrm{G}(n,p(n))$ 
a.a.s. is disconnected only if it contains an isolated vertex. 
An observation by Albert Atserias is the following:

\begin{theorem}
Let $c$ be a real constant such that 
$\Ln \mathrm{Pr}\big( G(n, c/n) 
\text{ is not }	3\text{-colorable }\big)> 0$.  
Then there is no FO graph property that explains 
non-$3$-colorability for $G(n, c/n)$.
\end{theorem}

The short proof of this theorem is as follows: 
It is a known fact that there are positive constants
$c_0\leq c_1$ such that $G(n, c/n)$ is a.a.s $3$-colorable
if $c<c_0$ and it is a.a.s non $3$-colorable if $c>c_1$ 
REFERENCES NEEDED.
Suppose $P$ is a FO graph property that implies
non-$3$-colorability. Then, because of this implication,
for all values of $c$
\[\Ln \mathrm{Pr}\big( G(n, c/n) 
\text{ satisfies } P \big) \leq \Ln \mathrm{Pr}\big( G(n, c/n) 
\text{ is not }	3\text{-colorable}\big).\]
In consequence the asymptotic probability that $G(n,c/n)$ 
satisfies $P$ is zero when $c<c_0$. By Lynch's theorem, if 
$P$ is definable in FO logic then this asymptotic probability
varies analytically with $c$. Using the fact that any analytic 
function that takes value zero in a non-empty interval must 
equal zero everywhere, we obtain that $G(n,c/n)$ a.a.s does 
not satisfy $P$ for any value of $c$. 
As a consequence the theorem follows.  \par

The aim of this work is to extend Lynch's result to arbitrary 
relational structures were the relations are subject to some
predetermined symmetry and anti-reflexivity axioms. This was 
originally motivated by an application to the study of random
$k$-CNF formulas. Since \cite{chvatal1992mick} it is known that 
for each $k$ there are constants $c_0,c_1$ such that a random 
$k$-CNF formula with $cn$ clauses over $n$ variables

%Besides this 
% 
%This question was originally posed in \cite{atserias2005definability}
%by Albert Atserias where the answer was given for various probability

\setcounter{section}{0}

@@

Given a positive natural number $n$, we will write
$[n]$ to denote the set ${1,2,\dots,n}$.\par
Given a set $S$ and a natural number $k\in \N$
we will use $\binom{S}{k}$ to denote the set of 
subsets of $S$ whose size is $k$. \par
Given numbers, $n,m\in \N$ with $m\leq n$ we denote by
$(n)_m$ the $m$-th falling factorial of $n$. \par
We will use the convention that over-lined variables
, like $\overline{x}$, denote ordered tuples of arbitrary length.
Given an ordered tuple $\overline{x}$
we define the number $len(\overline{x})$ as its length. 
Given a tuple $\overline{x}$ and an element $x$ the expression
$x\in \overline{x}$ means that $x$ appears as some coordinate
in $\overline{x}$. 
Given a map $f:X\rightarrow Y$ between two sets $X, Y$ and 
an ordered tuple $\overline{x}:=(x_1,\dots,x_a)\in X^*$ 
we define $f(\overline{x})\in Y^*$ as the tuple 
$(f(x_1),\dots,f(x_a))$.
Given two tuples $\overline{x},\overline{y}$
we write $\overline{x}^\smallfrown \overline{y}$ to denote their 
concatenation.
\par

Given a set
$S$ and a number $n\in \N$ with $n\leq |S|$ we define
$(S)_n$ as the subset of $S^n$ consisting of the $n$-tuples
whose coordinates are all different. 
We also define $S^*:=\bigcup_{n=0}^\infty S^n$ and
$(S)_*:=\bigcup_{n\leq |S|} (S)_n$. \par

@@

%\todo[inline]{Añadir referencias y trabajo relacionado con random SAT}

\subsection{Logical preliminaries}
We assume familiarity with first order logic (FO). We follow 
the convention that first order logic contains the equality symbol. 
Given a vocabulary $\sigma$ we will denote by $FO[\sigma]$ the set of 
first order formulas of vocabulary $\sigma$. We define the set of 
\textbf{free variables} of a formula as usual. 
Given a relation symbol $R\in \sigma$ we denote by $ar(R)$ the arity of $R$. 
Given a formula $\phi\in FO[\sigma]$ we will use the notation $\phi(\overline{y})$ 
to denote that $\overline{y}$ is a tuple of 
(different) variables that contains all free variables in $\phi$ and
none of its bounded variables, although it may contain variables
which not appear in $\phi$.
Formulas with no free variables are called \textbf{sentences} and 
formulas with no quantifiers are called \textbf{open formulas}.

@@

We define the class $\mathcal{C}$ as the 
class of $\sigma$-structures that
satisfy the 
following axioms: 
\begin{itemize}
	\item \textit{Symmetry axioms}: For each $R\in \sigma$ and
	each $g\in \Phi_R$:
	\[ \forall \overline{x}:=x_1,\dots, x_{ar(R)} \big(  R(\overline{x})
	\iff R(g\cdot\overline{x}) \big)    \]
	\item \textit{Anti-reflexivity axioms}: For each 
	$R\in \sigma$ and $\{i,j\}\in P_R$
	\[ \forall x_1,\dots, x_{ar(R)} 
	\big( (x_i=x_j) \implies \neg R(x_1,\dots, x_{a_s})
	\big)\]
\end{itemize}

@@

Given a set of vertices $U\subseteq V(G)$, 
we will denote by $G[U]$ the \textbf{hypergraph induced
by $G$ on $U$}. That is, $G[U]$ is an hypergraph
$H=(V(H),\{E(H)_R\}_{R\in \sigma})$ such that 
$V(H)=U$ and for any $R\in \sigma$ 
an edge $e\in E_R(G)$ belongs 
to $E_R(H)$ if and only if $V(e)\subset U$.
\par
%
%An homomorphism between two hypergraphs $G$ and $H$ 
%a map $f:V(G)\rightarrow V(H)$ that sends edges 
%from $G$ to edges in $H$ of the same color. That is,
%if vertices $v_1,\dots, v_{a_i}$ form an edge 
%$[v_1,\dots,v_{a_i}]\in E_i(G)$, then
%$[f(v_1),\dots,f(v_{a_i})]\in E_i(H)$.
%If $f$ is injective then it is called a monomorphism. 
%If $f$ is bijective and its inverse is also an homomorphism 
%between $H$ and $G$ then $f$ is called an isomorphism.\par
%
%The group of automorphisms $Aut(G)$ of an hypergraph $G$ is the group
%of isomorphisms between $G$ and itself.

%
%Given two hypergraphs $G$ and $H$, a copy of $H$ in $G$ is
%a sub-hypergraph $H_2\subseteq G$ isomorphic to $H$. The copy
%is called induced if $H_2$ is an induced sub-hypergraph.
%We will call a labeled copy
%of $H$ in $G$ to a monomorphism $f:H\rightarrow G$. 
%It is satisfied that the number
%of labeled copies of $H$ in $G$ is $\aut(H)$ times the number
%of copies of $H$ in $G$.

We define the \textbf{excess} $ex(G)$ of an hypergraph $G$ as the number
\[
ex(G):= \big(\sum_{R\in \sigma} (ar(R)-1)|E_R(G)|\big) - |V(G)|.  
\] 
That is, the excess of $G$ is its "weighted number of edges"
minus its number of vertices. \par
An hypergraph $G$ is \textbf{connected} if for any two vertices $v,u\in V(G)$
there is a sequence of edges $e_1,\dots, e_m\in E(G)$ such that
$v\in V(e_1), u\in V(e_m)$ and for each $i\in [m-1]$, 
$V(e_i)\cap V(e_{i+1})\neq \emptyset$. It holds that
$ex(G)\geq -1$ for any connected hypergraph.
\par
A connected hypergraph $G$ is a path between two of its 
vertices $v,u\in V(G)$ if $G$ 
does not contain any connected proper sub-hypergraph containing both $v,u$.
\par
A connected hypergraph $G$ is a \textbf{tree} if $ex(G)=-1$ and \textbf{dense} if $ex(G)>0$.
A connected hypergraph $G$ with $ex(G)\geq 0$ is called \textbf{saturated} 
if for any non-empty proper sub-hypergraph
$H\subset G$ it holds $ex(H)<ex(G)$. 
A connected hypergraph $G$ with $ex(G)=0$ is called \textbf{unicycle}. 
A saturated unicycle is called a \textbf{cycle}. \par
Given an hypergraph $G$ we define the following metric, $d$, over $V(G)$:
\[ d^G(u,v)= \min_{\substack{H \subset G\\ 
		H \text{ connected }\\
		u,v\in V(H)}} |E(H)| .\]
That is, the \textbf{distance} between $v$ and $u$ is the minimum number of
edges necessary to connect $v$ and $u$. 
If such number does not exist we define $d^G(u,v)=\infty$. 
When $G$ 
is understood or not relevant we will usually simply denote the 
distance by $d$ instead of $d^G$. Equivalently, the distance $d$
coincides with the usual one defined over the Gaifman graph of the structure 
$G$. The \textbf{diameter} of an hypergraph is the maximum distance between any 
two of its vertices. 
We extend naturally the distance $d$ to sets and tuples of
vertices, as usual. Given a vertex/set/tuple $X$ and a number
$r\in \N$ we define the \textbf{neighborhood}
$N^G(X;r)$, or simply $N(X;r)$ when $G$ is not relevant,
as the set of vertices $v$ such that $d^G(X,v)\leq r$.
\par

@@

Let $H$ be an hypergraph and let $V$ be a set. We define the
set of \textbf{copies of $H$ over $V$}, denoted as $Copies(H,V)$, 
as the set of hypergraphs 
$H^\prime$ such that
$V(H^\prime)\subset V$ and $H\simeq H^\prime$. Let 
$\chi$ be a $\Sigma$-coloring of $H$. 
As before, we define the set $Copies\big(
(H,\chi),\, \, V\big)$ as the set of $\Sigma$-hypergraphs
$(H^\prime,\chi^\prime)$ satisfying $V(H^\prime)\subset V$ and
$(H,\chi)\simeq (H^\prime,\chi^\prime)$. Let $\mathbb{H}$ be 
an isomorphism class of $\Sigma$-hypergraphs. Then the set
$Copies(\mathbb{H}, V)$ is defined as the set of $\Sigma$-hypergraphs
$(H^\prime,\chi^\prime)$ such that 
$V(H^\prime)\subset V$ and
$(H^\prime,\chi^\prime)\in \mathbb{H}$. 
Let $v\in V$ and $s\in \Sigma$. We define the
set $Copies\big(\mathbb{H}, V;\,\, (v,s)\big)$ 
as the set of $\Sigma$-hypergraphs
$(H^\prime,\chi^\prime)\in Copies(\mathbb{H}, V)$
that satisfy $v\in V(H^\prime)$ as well as
$\chi^\prime(v)=s$. \par

@@

For each $R\in \sigma$ let
$p_R$ be a real number between zero and one.
The random model $G^{\mathcal{C}}\Big(n,\InR{p}\Big)$ 
is the discrete probability space that
assigns to each hypergraph $G$ whose vertex
set $V(G)$ is $[n]$ the following probability:

@@

\begin{theorem} \label{thm:main}
	Let $\phi$ be a sentence in $FO[\sigma]$. Then
	the function
	%$F_\phi: [0,1]^{t_u}\times [0,\infty)^{t-t_u}
	$F_\phi: [0,\infty)^{|\sigma|}
	\rightarrow \R$ given by 
	\[
	\InR{\beta} \mapsto \Ln Pr\big( G_n\left(
	\InR{\beta}\right) \models \phi\big)
	\]
	is well defined and analytic. 
\end{theorem}

@@

\begin{definition}
Let $H$ be a connected hypergraph. If $ex(H)>0$, then it holds that
$H$ contains a unique maximal saturated sub-hypergraph $H^\prime$. 
Furthermore, $H^\prime$ satisfies $ex(H^\prime)=ex(H)$. If $ex(H)\leq 0$ the
set $H^\prime$ as the empty hypergraph. Given
$\overline{v}\in V(H)^*$ we define $Center(H,\overline{v})$ as the 
minimal connected sub-hypergraph in $H$ that contains both $H^\prime$
and the vertices in $\overline{v}$. 
If $H$ is a general hypergraph, not necessarily connected, 
we define $Center(H,\overline{v})$,
%	, resp. $Center(H,\overline{v};\, r)$, 
as the union, for all connected components $H^\prime \subset H$,
of the minimal connected sub-hypergraph of $H^\prime$ that
contains both $Center(H^\prime)$
%	, resp. $Center(H^\prime,\overline{v};\, r)$,
and the vertices in $\overline{v}$
that belong to $V(H^\prime)$.
\end{definition}

>>>>

\makeatletter
\newtheorem*{rep@theorem}{\rep@title}
\newcommand{\newreptheorem}[2]{%
	\newenvironment{rep#1}[1]{%
		\def\rep@title{#2 \ref{##1}}%
		\begin{rep@theorem}}%
		{\end{rep@theorem}}}
\makeatother

@@

\theoremstyle{definition}
\newtheorem{proposition}{Proposition}[section]
\newtheorem{fact}{Fact}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{obs}{Observation}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{propdef}{Proposition / Definition}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem*{theorem*}{Theorem}
\newreptheorem{theorem}{Theorem}
\newreptheorem{corollary}{Corollary}

\newenvironment{oneshot}[1]{\@begintheorem{#1}{\unskip}}{\@endtheorem}
\newcommand{\cc}{\mathfrak{c}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Hc}{\mathcal{H}}
\newcommand{\Lan}{\mathcal{L}}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\newcommand{\clist}{\mathfrak{c}_{1}, \cdots, \mathfrak{c}_m}
\newcommand{\morph}[1]{\sim_#1}
\newcommand{\vlst}[2]{#1_1,\dots, #1_{#2}}
\newcommand{\gnp}{G(n,\beta_1/n^{a_1-1}, \dots,\beta_l/n^{a_l-1})}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\ehr}{\textsc{Ehr}}
\newcommand{\PR}[1]{\mathrm{Pr}\left(#1\right)}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\InR}[1]{\left\{ #1_R \right\}_{R\in \sigma}}
\newcommand{\sep}{\noindent\rule{2cm}{0.4pt}}
\newcommand{\aut}{\mathrm{aut}}

@@

We say that a sequence of random structures $\{G_n\}_n$ satisfies a limit
law with respect to some logical language $L$ if for any property $P$ expressible in
$L$ the probability that $G_n$ satisfies $P$ tends to some limit as $n$ goes to $\infty$.
If that limit takes only the values zero and one then we say that $\{G_n\}_n$ satisfies
a zero-one law with respect to $L$.  
Convergence and zero one laws have been extensively studied with regards to 
random binomial graphs $G(n,p)$.
The seminal theorem on this topic, due to Fagin \cite{fagin1976probabilities}
and Glebskii et al. \cite{glebskii1969range} independently, 
concerns general relational structures.
When applied to graphs it states that if $p$ is a fixed probability
then $G(n,p)$ satisfies a zero-one law with respect to the first order
(FO) language of graphs. This zero-one law was later extended by Shelah and Spencer
in \cite{shelah1988zero}. There it is proven, among other results,
that if $p:=p(n)$ was a decreasing function on
$n$ of the form $n^{-\alpha}$, and $\alpha\in (0,1)$ is irrational
then $G(n,p(n))$ obeys a zero-one law with respect to FO logic. Moreover, 
it is also proven that if $\alpha\in (0,1)$ is rational then $G(n,p(n))$ 
does not obey a convergence law. This was further extended by Lynch
in \cite{lynch1992probabilities}, where it is shown that in the
case where the expected number of edges is linear, when $p(n)\sim \beta/n$
for some $\beta\in (0,\infty)$, then $G(n,p(n))$ satisfies a limit law with
respect to FO logic. The following is a restatement of the main theorem
in that article.

\begin{theorem*}
	Let $p(n)\sim \beta/n$. For any FO sentence $\phi$, the function
	$F_\phi: (0,\infty)\rightarrow [0,1]$ given by 
	\[ F_\phi(\beta) = \Ln \mathrm{Pr}\left( \mathrm{G}(n,p(n))
	\text{ satisfies } \phi   \right) \]
	is well defined and is given by an expression with parameter $\beta$ built
	using rational constants, addition, multiplication and exponentiation with base $e$.
\end{theorem*}

A relevant aspect of this result is that the limit probability of any FO
property in $G(n,p(n))$ when $p(n)\sim \beta/n$ varies analytically with $\beta$.
A consequence of this is that in this conditions no FO property can "capture" sudden changes 
in the structure of $G(n,p(n))$. \par
It was left open at the end of \cite{lynch1992probabilities} whether the convergence law obeyed
by $G(n,\beta/n)$ could be generalized for other random models of relational structures
that contain relations of degree greater than $2$.
A result in this regard was obtained in \cite{salvadorbrasil}, 
among other zero-one and convergence laws.
There it is considered $G^d(n,p)$ the random model of $d$-uniform hypergraphs where
each $d$-edge is added to a set of $n$ labeled vertices with probability $p$ independently.
It is shown that when $p(n)\sim \beta/n^{d-1}$ -i.e., when the expected number of
edges is linear- then $G^d(n,p(n))$ obeys a convergence law with respect of the FO
language of $d$-uniform hypergraphs. With little additional work it can be shown that
in this conditions the limit probability of any FO property of $G^d(n,p(n))$ varies 
analytically with $\beta$. \par
We aim to extend this last result to arbitrary relational structures
to whose relations we can impose symmetry and anti-reflexivity constraints. 
This generalization is motivated by an application to the problem of random 
satisfiability. We continue the study started by Atserias in \cite{atserias2005definability}
with regards to the definability in first order logic of certificates for unsatisfiability
that hold for typical unsatisfiable formulas. There it is considered a random model for
$3$-CNF formulas where each possible clause over $n$ variables is added with probability $p$ 
independently. This way the expected number of clauses is $\Theta(n^3p)$ as $n$ grows.
The main result of that article states that when the expected number of clauses is $\Theta(n^{2+\alpha})$
for any arbitrary $\alpha>0$ then there exists some FO property of $3$-CNF formulas that implies
unsatisfiability and asymptotically almost surely (a.a.s) holds for unsatisfiable formulas. 
Furthermore, if the expected number of clauses is $\Theta(n^{2-\alpha})$ for any irrational
$\alpha>0$ then no FO property that implies unsatisfiability holds a.a.s
for unsatisfiable formulas.

@@

Given a positive natural number $n$, we will write
$[n]$ to denote the set ${1,2,\dots,n}$.\par
Given numbers, $n,m\in \N$ with $m\leq n$ we denote by
$(n)_m$ the $m$-th falling factorial of $n$. \par
Given a set $S$ and a natural number $k\in \N$
we will use $\binom{S}{k}$ to denote the set of 
subsets of $S$ whose size is $k$.Given a set
$S$ and a number $n\in \N$ with $n\leq |S|$ we define
$(S)_n$ as the subset of $S^n$ consisting of the $n$-tuples
whose coordinates are all different. 
We also define $S^*:=\bigcup_{n=0}^\infty S^n$ and
$(S)_*:=\bigcup_{n\leq |S|} (S)_n$. \par

We will use the convention that over-lined variables
, like $\overline{x}$, denote ordered tuples of arbitrary length.
Given an ordered tuple $\overline{x}$
we define the number $len(\overline{x})$ as its length. 
Given a tuple $\overline{x}$ and an element $x$ the expression
$x\in \overline{x}$ means that $x$ appears as some coordinate
in $\overline{x}$. 
Given a map $f:X\rightarrow Y$ between two sets $X, Y$ and 
an ordered tuple $\overline{x}:=(x_1,\dots,x_a)\in X^*$ 
we define $f(\overline{x})\in Y^*$ as the tuple 
$(f(x_1),\dots,f(x_a))$.
Given two tuples $\overline{x},\overline{y}$
we write $\overline{x}^\smallfrown \overline{y}$ to denote their 
concatenation.
\par

@@

%%\todo[inline]{Añadir referencias y trabajo relacionado con random SAT}
%

\subsection{Logical preliminaries}
We assume familiarity with first order logic (FO). We follow 
the convention that first order logic contains the equality symbol. 
Given a vocabulary $\sigma$ we will denote by $FO[\sigma]$ the set of 
first order formulas of vocabulary $\sigma$. We define the set of 
\textbf{free variables} of a formula as usual. 
Given a relation symbol $R\in \sigma$ we denote by $ar(R)$ the arity of $R$. 
Given a formula $\phi\in FO[\sigma]$ we will use the notation $\phi(\overline{y})$ 
to express that $\overline{y}$ is a tuple of 
(different) variables which contains all free variables in $\phi$ and
none of its bounded variables, but it may contain variables
which do not appear in $\phi$.
Formulas with no free variables are called \textbf{sentences} and 
formulas with no quantifiers are called \textbf{open formulas}.

@@

We define the class $\mathcal{C}$ as the 
class of $\sigma$-structures that
satisfy the 
following axioms: 
\begin{itemize}
	\item \textit{Symmetry axioms}: For each $R\in \sigma$ and
	each $g\in \Phi_R$:
	\[ \forall \overline{x}:=x_1,\dots, x_{ar(R)} \left(  R(\overline{x})
	\iff R(g\cdot\overline{x}) \right)    \]
	\item \textit{Anti-reflexivity axioms}: For each 
	$R\in \sigma$ and $\{i,j\}\in P_R$
	\[ \forall x_1,\dots, x_{ar(R)} 
	\left( (x_i=x_j) \implies \neg R(x_1,\dots, x_{a_s})
	\right)\]
\end{itemize}

@@

Given a set of vertices $U\subseteq V(G)$, 
we will denote by $G[U]$ the \textbf{hypergraph induced
by $G$ on $U$}. That is, $G[U]$ is an hypergraph
$H=(V(H),\{E(H)_R\}_{R\in \sigma})$ such that 
$V(H)=U$ and for any $R\in \sigma$ 
an edge $e\in E_R(G)$ belongs 
to $E_R(H)$ if and only if $V(e)\subset U$.
\par

We define the \textbf{excess} $ex(G)$ of an hypergraph $G$ as the number
\[
ex(G):= \left(\sum_{R\in \sigma} (ar(R)-1)|E_R(G)|\right) - |V(G)|.  
\] 
That is, the excess of $G$ is its "weighted number of edges"
minus its number of vertices. \par
An hypergraph $G$ is \textbf{connected} if for any two vertices $v,u\in V(G)$
there is a sequence of edges $e_1,\dots, e_m\in E(G)$ such that
$v\in V(e_1), u\in V(e_m)$ and for each $i\in [m-1]$, 
$V(e_i)\cap V(e_{i+1})\neq \emptyset$. It holds that
$ex(G)\geq -1$ for any connected hypergraph.
\par
Given an hypergraph $G$ we define the following metric, $d$, over $V(G)$:
\[ d^G(u,v)= \min_{\substack{H \subset G\\ 
		H \text{ connected }\\
		u,v\in V(H)}} |E(H)| .\]
That is, the \textbf{distance} between $v$ and $u$ is the minimum number of
edges necessary to connect $v$ and $u$. 
If such number does not exist we define $d^G(u,v)=\infty$. 
When $G$ 
is understood or not relevant we will usually simply denote the 
distance by $d$ instead of $d^G$. Equivalently, the distance $d$
coincides with the usual one defined over the Gaifman graph of the structure 
$G$. The \textbf{diameter} of an hypergraph is the maximum distance between any 
two of its vertices. 
We extend naturally the distance $d$ to sets and tuples of
vertices, as usual. Given a vertex/set/tuple $X$ and a number
$r\in \N$ we define the \textbf{neighborhood}
$N^G(X;r)$, or simply $N(X;r)$ when $G$ is not relevant,
as the set of vertices $v$ such that $d^G(X,v)\leq r$.
\par
A connected hypergraph $G$ is a path between two of its 
vertices $v,u\in V(G)$ if $G$ 
does not contain any connected proper sub-hypergraph containing both $v,u$.
\par
A connected hypergraph $G$ is a \textbf{tree} if $ex(G)=-1$ and \textbf{dense} if $ex(G)>0$.
An hypergraph is called $r$-\textbf{sparse} if it does not contain any dense sub-hypergraph $H$
such that $diam(H)\leq r$.
A connected hypergraph $G$ with $ex(G)\geq 0$ is called \textbf{saturated} 
if for any non-empty proper sub-hypergraph
$H\subset G$ it holds $ex(H)<ex(G)$. 
A connected hypergraph $G$ with $ex(G)=0$ is called \textbf{unicycle}. 
A saturated unicycle is called a \textbf{cycle}. \par
A \textbf{rooted tree} $(T,v)$ is a tree $T$ with a 
distinguished vertex $v\in V(T)$ called its \textbf{root}.
We will usually omit the root when it is not relevant and 
write just $T$ instead of $(T,v)$. The
\textbf{initial edges} of a rooted tree $(T,v)$ are 
the edges in $T$ that contain $v$. 
We define the radius of a rooted tree
as the maximum distance between its root
and any other vertex.
\par

@@

Let $H$ be an hypergraph and let $V$ be a set. We define the
set of \textbf{copies of $H$ over $V$}, denoted as $Copies(H,V)$, 
as the set of hypergraphs 
$H^\prime$ such that
$V(H^\prime)\subset V$ and $H\simeq H^\prime$. Let 
$\chi$ be a $\Sigma$-coloring of $H$. 
As before, we define the set $Copies\left(
(H,\chi),\, \, V\right)$ as the set of $\Sigma$-hypergraphs
$(H^\prime,\chi^\prime)$ satisfying $V(H^\prime)\subset V$ and
$(H,\chi)\simeq (H^\prime,\chi^\prime)$. Let $\mathbb{H}$ be 
an isomorphism class of $\Sigma$-hypergraphs. Then the set
$Copies(\mathbb{H}, V)$ is defined as the set of $\Sigma$-hypergraphs
$(H^\prime,\chi^\prime)$ such that 
$V(H^\prime)\subset V$ and
$(H^\prime,\chi^\prime)\in \mathbb{H}$. 
Let $v\in V$ and $s\in \Sigma$. We define the
set $Copies\left(\mathbb{H}, V;\,\, (v,s)\right)$ 
as the set of $\Sigma$-hypergraphs
$(H^\prime,\chi^\prime)\in Copies(\mathbb{H}, V)$
that satisfy $v\in V(H^\prime)$ as well as
$\chi^\prime(v)=s$. \par

@@

For each $R\in \sigma$ let
$p_R$ be a real number between zero and one.
The random model $G^{\mathcal{C}}\left(n,\InR{p}\right)$ 
is the discrete probability space that
assigns to each hypergraph $G$ whose vertex
set $V(G)$ is $[n]$ the following probability:

@@

\begin{theorem} \label{thm:main}
	Let $\phi$ be a sentence in $FO[\sigma]$. Then
	the function
	%$F_\phi: [0,1]^{t_u}\times (0,\infty)^{t-t_u}
	$F_\phi: (0,\infty)^{|\sigma|}
	\rightarrow \R$ given by 
	\[
	\InR{\beta} \mapsto \Ln Pr\left( G_n\left(
	\InR{\beta}\right) \models \phi\right)
	\]
	is well defined and analytic. 
\end{theorem}

@@

\begin{definition} 
Let $H$ be a connected hypergraph. Then it holds that
$H$ contains a unique maximal saturated sub-hypergraph $H^\prime$
and it satisfies $ex(H^\prime)=ex(H)$ if $ex(H)\geq 0$ and $H^\prime=\emptyset$ 
if $H$ is a tree. Given
$\overline{v}\in V(H)^*$ we define $Center(H,\overline{v})$ as the 
minimal connected sub-hypergraph in $H$ that contains both $H^\prime$
and the vertices in $\overline{v}$. 
If $H$ is not connected 
we define $Center(H,\overline{v})$,
%	, resp. $Center(H,\overline{v};\, r)$, 
as the union of $Center(H^{\prime\prime},\overline{u})$
for all connected components $H^{\prime\prime} \subset H$,
where $\overline{u}\in V(H)^*$ contains exactly the vertices
in $\overline{v}$ belonging to $V(H^{\prime\prime})$. When $\overline{v}$
is empty we simply write $Center(H)$.
\end{definition}

\begin{definition}
Let $H$ be an hypergraph, $\overline{v}\in V(H)^*$
and $r\in \N$. Let $X$ be the set of vertices $v\in V(H)$
that either belong to $\overline{v}$ or belong 
to some saturated sub-hypergraph of $H$ with
diameter at most $2r+1$. 
We define $Core(H, \overline{v};r)$ as $N(X;r)$. If
$\overline{v}$ is empty we write $Core(H;r)$.
We say that $H$ is \textbf{$r$-simple} if all connected components of 
$Core(H;r)$ are unicycles. 	
\end{definition}
